\section{Events}

This section enumerates simulation-managed events that were dispatched on virtual CPUs.
In addition to a program, each genome contained an array of 64-bit tags --- one for each event.
When an event's criteria was met in the simulation, the genome's corresponding tag was used to dispatch a module in the program and launch a core executing that module.

All events are also exposed to the cell as a corresponding input sensor.
The state of the event (0 for false, 1 for true) is stored in the sensor prior to virtual CPU execution.
In fact, events are triggered based on the reading of the sensor register (not by re-reading the underlying simulation state).
This means that experimental perturbations that perturb sensor input also disrupted event-handling, allowing the state interface complexity metric to measure both event-driven and sensor-based behaviors.

The source code controlling events can be found at \url{https://github.com/mmore500/dishtiny/tree/prq49/include/dish2/events} and \url{https://github.com/mmore500/dishtiny/blob/prq49/include/dish2/services/InterpretedIntrospectiveStateRefreshService.hpp}.

\subsection{Always}

This event is always dispatched.

\subsection{Is Child Cell Of}

Is this cell a daughter cell of the cardinal's neighbor?
Triggered if this cell was spawned from the cardinal's neighbor and its cell is younger than the neighbor.

\subsection{Is Child Group Of (0 thru $L-1$)}

Is this cell's kin group a daughter group of the cardinal's neighbor cell's kin group?
Triggered if a cell's kin group ancestor ID(s) are equal to the cardinal's neighbor's current kin group ID(s).

\subsection{Is Newborn}

This event is dispatched once when a cell is first born.
Triggered if cell age is less than frequency at which events are launched.

\subsection{Is Parent Cell Of}

Is this cardinal's cell the parent cell of the cardinal's neighbor?
Triggered if neighbor was spawned from cell and cell age is greater than neighbor age.

\subsection{Kin Group Match (0 thru $L-1$)}

Is this cell part of the same kin group as the cardinal's neighbor?
Triggered if a cell's kin group ID(s) are equal to the cardinal's neighbors' current kin group ID(s).

\subsection{Kin Group Mismatch (0 thru $L-1$)}

Is this cell part of a different kin group from the cardinal's neighbor cell?
Triggered if a cell's kin group ID(s) are not equal to the cardinal's neighbors' current kin group ID(s).

\subsection{Kin Group will Expire (0 thru $L-1$)}

Triggered if kin group age is greater than 80\% of the kin group expiration duration.
(Depending on experiment configuration, the group may be force-fragmented after expiration.)

\subsection{Kin Group will not Expire (0 thru $L-1$)}

Triggered if kin group age is less than or equal to than 80\% of the kin group expiration duration.

\subsection{Neighbor Apoptosis}

Triggered if the most recent cell death in the cardinal's neighbor tile was apoptosis.

\subsection{Neighbor Fragmented}

Triggered if the most recent cell death in the cardinal's neighbor tile was fragmentation.

\subsection{Neighbor Is Alive}

Triggered if a cardinal's neighbor tile is occupied by a live cell.

\subsection{Neighbor Is Newborn}

Triggered once for each time a newborn spawns into the cardinal's neighboring tile.
Triggered if the cardinal's neighbor's age is less than the frequency at which events are launched.

\subsection{Neighbor Is Not Alive}

Triggered if the cardinal's neighboring tile is not occupied.

\subsection{Neighbor Kin Group Will Expire (0 thru $L-1$)}

Triggered if the cardinal's cell neighbor's kin group age is less than or equal to 80\% of the kin group expiration duration.

\subsection{Neighbor Optimum Quorum Exceeded}

Triggered if the cardinal's cell neighbor's number of known quorum bits exceed the target quorum count.

\subsection{Optimum Quorum Exceeded (0 thru $L-1$)}

Triggered if the cell's number of known quorum bits exceed the target quorum count.

\subsection{Optimum Quorum Not Exceeded (0 thru $L-1$)}

Triggered if the cell's number of known quorum bits is less than or equal to than the target quorum count.

\subsection{Parent Fragmented}

Triggered if a cell's parent died from fragmentation.
That is, if the last cause of death on the current tile was fragmentation.

\subsection{Phylogenetic Root Match}

Does this cell descend from the same originally-generated genome as its neighbor?
Triggered if a cardinal's cell's root ID is equal to that cardinal's neighbor cell's root ID.

\subsection{Phylogenetic Root Mismatch}

Does this cell and its neighbor descend from a different originally-generated genomes?
Triggered if a cardinal's cell's root ID is not equal to that cardinal's neighbor cell's Root ID.

\subsection{Poorer Than Neighbor}

Does this cell have less resource stockpiled than its neighbor?
Triggered if a cardinal's cell has less resource than that cardinal's neighbor cell.

\subsection{Received Resource From}

Triggered if a cardinal's cell has received resource from that cardinal's neighbor cell.

\subsection{Richer Than Neighbor}

Does this cell have more resource stockpiled than its neighbor?
Triggered if a cardinal's cell has more resource in its stockpile than that cardinal's neighbor cell.

\subsection{Stockpile Depleted}

Is this cell's stockpile empty?
Triggered if a cell's stockpile is less than twice the base harvest rate.

\subsection{Stockpile Fecund}

Does this cell have enough stockpiled resource to fund cellular reproduction?
Triggered if a cell's stockpile is greater than 1.0.
