% create command to draw state tables
\newcommand{\cellsimservicedef}[2]{
    \begin{tabular}{|
        >{\columncolor[HTML]{C0C0C0}}l |l|}
        \hline
        Frequency & every #1 update(s) \\ \hline
    \end{tabular} \\
}

\section{Cellular Simulation Services}

Simulation logic is applied to each cell through a collection of distinct functors, referred to as services.

All services specified to run on a particular update are applied in sequence to a single cell.
(Some services run only every $n$th update.)
Then, to another randomly-chosen cell in a \texttt{thread\_local} population, and another until the entire population has been updated.

See \url{https://github.com/mmore500/dishtiny/tree/prq49/include/dish2/services} for source code implementing these services.

\subsection{Decay to Baseline Service}

\cellsimservicedef{32}

Decays a cell's global regulators, resets its controller-mapped peripheral states, and resets its transient NOP states.

\subsection{Running Log Purge Service}

\cellsimservicedef{64}

Purges a cell's running logs.
(Only affects data collection, not simulation logic.)

\subsection{Controller Mapped State Noise Service}

\cellsimservicedef{8}

Given a non-zero controller-mapped state defect rate, picks a random number $n$ from a Poisson distribution parameterized by \texttt{CONTROLLER\_MAPPED\_STATE\_DEFECT\_RATE}.
Then, it introduces $n$ defects to a cell's writable state.
Half of these defects zero out the state and half randomize it.

\subsection{Interpreted Introspective State Refresh Service}

\cellsimservicedef{i}

Refreshes the interpreted introspective state of a cell.

\subsection{Extrospective State Exchange Service}

\cellsimservicedef{1}

Used for experimental manipulations testing the fitness effect of extrospective state.
(Not part of core simulation logic.)

\subsection{Extrospective State Rotate Service}

\cellsimservicedef{1}

Used for experimental manipulations testing the fitness effect of extrospective state.
(Not part of core simulation logic.)

\subsection{Introspective State Exchange Service}

\cellsimservicedef{1}

Used for experimental manipulations testing the fitness effect of introspective state.
(Not part of core simulation logic.)

\subsection{Introspective State Rotate Service}

\cellsimservicedef{1}

Used for experimental manipulations testing the fitness effect of introspective state.
(Not part of core simulation logic.)

\subsection{CPU Execution Service}

\cellsimservicedef{1}

Executes a cell's genome on its cardinals' virtual CPUs for \texttt{HARDWARE\_EXECUTION\_CYCLES} virtual cycles.
The order of cardinal evaluation is randomized.
This is repeated \texttt{HARDWARE\_EXECUTION\_ROUNDS} times.

\subsection{Event Launching Service}

\cellsimservicedef{8}

Dispatches environmentally-managed events for each cardinal.

\subsection{Introspective State Rotate Restore Service}

\cellsimservicedef{1}

Used for experimental manipulations testing the fitness effect of introspective state.
(Not part of core simulation logic.)

\subsection{Introspective State Exchange Restore Service}

\cellsimservicedef{1}

Used for experimental manipulations testing the fitness effect of introspective state.
(Not part of core simulation logic.)

\subsection{Extrospective State Rotate Restore Service}

\cellsimservicedef{1}

Used for experimental manipulations testing the fitness effect of extrospective state.
(Not part of core simulation logic.)

\subsection{Extrospective State Exchange Restore Service}

\cellsimservicedef{1}

Used for experimental manipulations testing the fitness effect of extrospective state.
(Not part of core simulation logic.)

\subsection{Writable State Exchange Service}

\cellsimservicedef{1}

Used for experimental manipulations testing the fitness effect of writable state.
(Not part of core simulation logic.)

\subsection{Writable State Rotate Service}

\cellsimservicedef{1}

Used for experimental manipulations testing the fitness effect of writable state.
(Not part of core simulation logic.)

\subsection{Birth Setup Service}

\cellsimservicedef{16}

Births a new cell into the current cell.

This occurs by first iterating through the cell's cardinals' birth request inputs in random order.
While the cell's resource stockpile is greater than the \texttt{SPAWN\_DEFENSE\_COST}, the requests are ignored and the stockpile depleted by that cost.
The first request that cannot be defended against is then acted upon.
The current cell's death routine is called, the old genome is replaced by the incoming genome, and the cell's make-alive routine is called.

\subsection{Cell Age Service}

\cellsimservicedef{1}

Advances the cell age introspective state and refreshes kin group age introspective state.

\subsection{Collective Harvesting Service}

\cellsimservicedef{4}

Calculates the total amount of resource collectively harvested to this cell by the cell's kin group.
This amount increases with quorum count and saturates at \texttt{OPTIMAL\_QUORUM\_COUNT}.
Adds the harvested amount to the cell's resource stockpile.

\subsection{Collective Resource Decay Service}

\cellsimservicedef{1}

If the cell's quorum count exceeds \texttt{OPTIMAL\_QUORUM\_COUNT}, applies multiplicative decay to the cell's resource stockpile.
This effect strengthens exponentially with excess cell quorum count.

\subsection{Conduit Flush Service}

\cellsimservicedef{16}

Flushes each cardinals' inter-process and inter-thread output conduits.

\subsection{Inter Message Launching Service}

\cellsimservicedef{8}

Launches new virtual cores to process incoming inter-cell messages.

\subsection{Inter Message Purging Service}

\cellsimservicedef{8}

Purges excess incoming inter-cell messages that couldn't be handled due to virtual core availability.

\subsection{Intra Message Launching Service}

\cellsimservicedef{8}

Launches new virtual cores to process incoming messages from co-cardinals.

\subsection{Message Counter Clear Service}

\cellsimservicedef{16}

Intermittently resets introspective message count state.

\subsection{Quorum Service}

\cellsimservicedef{1}

Performs distributed estimation of kin group size by simulation.

Each cell has a single randomly-chosen index set within a fixed-length bitstring.
(Depending in parameter settings, some cells may may have index set --- all positions within the bitstring are zeroed out.)

Broadcasts bits known to be set are to all neighbor cells within the same kin group.
Incoming bitstrings from neighbors are ORed with known bits.

The original neighbor each non-self bit was first learned from is recorded alongside that bit.
If that neighbor no longer broadcasts that bit, it is erased from the cell's known bits.

Updates latest quorum count into introspective state.

This scheme is replicated independently for each kin group level simulated.

\subsection{Resource Decay Service}

\cellsimservicedef{1}

Decays cell resource stockpile multiplicatively by \texttt{RESOURCE\_DECAY} constant.

\subsection{Resource Harvesting Service}

\cellsimservicedef{1}

Adds a constant amount to cell's resource stockpile.

\subsection{Resource Receiving Service}

\cellsimservicedef{4}

Calculates total amount of resource received across every cardinal, and then adds that total to resource stockpile.

If the cell is not alive, it instead refunds all received resources back to each sending cell.

\subsection{Resource Sending Service}

\cellsimservicedef{1}

Based on writable state within each cardinal, calculates and dispatches resource that should be shared to each neighbor cell.

\subsection{Spawn Sending Service}

\cellsimservicedef{16}

If available resource is greater than or equal to 1.0, iterates randomly through every cardinal to determine whether it requested to spawn and has not arrested spawning.
Then, one of these requests is dispatched at random and stockpile is decreased by one.

\subsection{State Input Jump Service}

\cellsimservicedef{8}

Pulls a fresh copy of each neighbor cardinal's current readable state.

\subsection{State Output Put Service}

\cellsimservicedef{8}

Dispatches a copy of each cardinal's current readable state to each cardinal's neighbor cell.

\subsection{Epoch Advance Service}

\cellsimservicedef{8}

The cell's current-known epoch count is advanced by one then set to the maximum of the cell's current-known epoch count and neighbor cells' current-known epoch count.

\subsection{Writable State Rotate Restore Service}

\cellsimservicedef{1}

Used for experimental manipulations testing the fitness effect of writable state.
(Not part of core simulation logic.)

\subsection{Writable State Exchange Restore Service}

\cellsimservicedef{1}

Used for experimental manipulations testing the fitness effect of writable state.
(Not part of core simulation logic.)

\subsection{Group Expiration Service}

\cellsimservicedef{64}

As group age exceeds \texttt{GROUP\_EXPIRATION\_DURATIONS}, with increasing probability fragments cell from its kin group.
This process kills the cell and replaces it in place with a daughter without kin ID commonality.

\subsection{Apoptosis Service}

\cellsimservicedef{16}

If any cardinals have requested apoptosis, do death routine on the cell.
