\section{Events}

See \url{https://github.com/mmore500/dishtiny/tree/prq49/include/dish2/events} and \url{https://github.com/mmore500/dishtiny/blob/prq49/include/dish2/services/InterpretedIntrospectiveStateRefreshService.hpp}.

\subsection{Always}

Always true.

\subsection{Is Child Cell Of}

Whether cell was spawned from a neighbor and cell age is less than neighbor age.

\subsection{Is Child Group Of (0 thru $L-1$)}

Whether its ancestor ID is equal to a cell neighbor's IDs.

\subsection{Is Newborn}

Whether cell age is less than frequency at which events are launched.

\subsection{Is Parent Cell Of}

Whether neighbor was spawned from cell and cell age is greather than neighbor age.

\subsection{Kin Group Match (0 thru $L-1$)}

\subsection{Kin Group Mismatch (0 thru $L-1$)}

\subsection{Kin Group will Expire (0 thru $L-1$)}

Whether Kin Group age is greater than 80\% of the Kin Group Expiration duration.

\subsection{Kin Group will not Expire (0 thru $L-1$)}

Whether Kin Group age is less than or equal to than 80\% of the Kin Group Expiration duration.

\subsection{Neighbor Apoptosis}

Whether a cell neighbor died of apoptosis.

\subsection{Neighbor Fragmented}

Whether a cell neighbor died of fragmentation.

\subsection{Neighbor Is Alive}

Whether a cell neighbor is alive.

\subsection{Neighbor Is Newborn}

Whether a cell neighbor's age is less than the frequency at which events are launched.

\subsection{Neighbor Is Not Alive}

Whether a cell's neighbor has undergone apoptosis.

\subsection{Neighbor Kin Group Will Expire (0 thru $L-1$)}

Whether a cell's neighbor's Kin Group age is less than or equal to 80\% of the Kin Group Expiration duration.

\subsection{Neighbor Optimum Quorum Exceeded}

Whether a cell's neighbor's known number of quorum bits are greater than the optimum quorum count.

\subsection{Optimum Quorum Exceeded (0 thru $L-1$)}

Whether a cell's known number of quorum bits are greater than the optimum quorum count.

\subsection{Optimum Quorum Not Exceeded (0 thru $L-1$)}

Whether a cell's known number of quorum bits are less than or equal to than the optimum quorum count.

\subsection{Parent Fragmented}

Whether a cell's parent died from fragmentation.

\subsection{Phylogenetic Root Match}

Whether a cell's Root ID is equal to a cell's neighbor's Root ID. This means they originate from the same phylogeny.

\subsection{Phylogenetic Root Mismatch}

Whether a cell's Root ID is not equal to a cell's neighbor's Root ID.

\subsection{Poorer Than Neighbor}

Whether a cell has less resource than a cell neighbor.

\subsection{Received Resource From}

Whether a cell has recieved resource from that cell neighbor.

\subsection{Richer Than Neighbor}

Whether a cell has more resource than a cell neighbor.

\subsection{Stockpile Depleted}

Whether a cell's stockpile is less than twice the base harvest rate.

\subsection{Stockpile Fecund}

Whether a cell's stockpile is greater than one.
