\section{Introspective State}

In addition to raw introspective state, all events have interpreted introspective state stored here.

See \url{https://github.com/mmore500/dishtiny/tree/prq49/include/dish2/peripheral/readable_state/introspective_state}.

\subsection{Is Child Cell Of}

\texttt{char}

1 if Is Child Cell Of event should be triggered, 0 otherwise.
See Is Child Cell Of event. (crossreference)

\subsection{Is Child Group Of (0 thru $L-1$)}

\texttt{char}

\subsection{Is Newborn}

\begin{tabular}{|
>{\columncolor[HTML]{C0C0C0}}l |l|}
\hline
Type   & \texttt{char}                                            \\ \hline
Values & 0 (\texttt{false}) or 1 (\texttt{true}) \\ \hline
\end{tabular}

Is the cell age less than \texttt{cfg.EVENT\_LAUNCHING\_SERVICE\_FREQUENCY}?

\subsection{Is Parent Cell Of}

This state is \texttt{char} type and stores the value 0 or 1.

Is this cell the parent cell of this cardinal's neighbor?

\subsection{Kin Group Match (0 thru $L-1$)}

\texttt{char}

\subsection{Kin Group Mismatch (0 thru $L-1$)}

\texttt{char}

\subsection{Kin Group will Expire (0 thru $L-1$)}

\texttt{char}

\subsection{Kin Group will not Expire (0 thru $L-1$)}

\texttt{char}

\subsection{Neighbor Apoptosis}

Whether neighbor has died due to apoptosis.

\texttt{char}

\subsection{Neighbor Fragmented}

Whether neighbor has died due to fragmentation.

\texttt{char}

\subsection{Neighbor Is Alive}

Whether neighbor is alive.

\texttt{char}

\subsection{Neighbor Is Newborn}

Whether neighbor age is less than frequency at which events are launched.

\texttt{char}

\subsection{Neighbor Is Not Alive}

Whether neighbor is not alive.

\texttt{char}

\subsection{Neighbor Kin Group Will Expire (0 thru $L-1$)}

\texttt{char}

Whether Kin Group age is greater than 80\% of the Kin Group Expiration duration.

\subsection{Neighbor Optimum Quorum Exceeded}

\texttt{char}

\subsection{Optimum Quorum Exceeded (0 thru $L-1$)}

\texttt{char}

Whether a cell's known number of quorum bits are greater than the optimum quorum count.

\subsection{Optimum Quorum Not Exceeded (0 thru $L-1$)}

\texttt{char}

Whether a cell's known number of quorum bits are less than or equal to than the optimum quorum count.

\subsection{Parent Fragmented}

\texttt{char}

Whether a cell's parent died from fragmentation.

\subsection{Phylogenetic Root Match}

\texttt{char}

Whether a cell's Root ID is equal to a cell's neighbor's Root ID. This means they originate from the same phylogeny.

\subsection{Phylogenetic Root Mismatch}

\texttt{char}

Whether a cell's Root ID is not equal to a cell's neighbor's Root ID.

\subsection{Poorer Than Neighbor}

\texttt{char}

Whether a cell has less resource than a cell neighbor.

\subsection{Received Resource From}

\texttt{char}

Whether a cell has recieved resource from that cell neighbor.

\subsection{Richer Than Neighbor}

\texttt{char}

Whether a cell has more resource than a cell neighbor.

\subsection{Stockpile Depleted}

\texttt{char}

Whether a cell's stockpile is less than twice the base harvest rate.

\subsection{Stockpile Fecund}

\texttt{char}

Whether a cell's stockpile is greater than one.

\subsection{Cell Age}

\texttt{char}

Updates elapsed since cell was born.

\subsection{Epoch}

\texttt{uint64\_t}

Updates elapsed since start of simulation.

\subsection{Incoming Inter Message Counter}

\texttt{uint64\_t}

Counter of incoming messages from other cells.

\subsection{Incoming Intra Message Counter}

\texttt{uint64\_t}

Counter of incoming messages from other cardinals within the cell.

\subsection{Is Alive}

\texttt{char}

Whether the cell is alive.

\subsection{Kin Group Age ($L\times$)}

\texttt{uint64\_t}

Number of updates since Kin Group formation.

\subsection{Kin Group ID Ancestor View ($L\times$)}

\texttt{uint64\_t}

Kin Group ID of the cell's ancestor.

\subsection{Kin Group ID View ($L\times$)}

\texttt{char}

Kin Group ID of this cell.

\subsection{Most Recent Cause of Death}

\texttt{char}
(see enum/CauseOfDeath)

\subsection{Num Known Quorum Bits ($L\times$)}

\texttt{uint64\_t}

\subsection{Phylogenetic Root View}

\texttt{uint64\_t}

\subsection{Received Resource From}

\texttt{float}

\subsection{Resource Stockpile}

\texttt{float}

Amount of resource this cell has.

\subsection{Spawn Count}

\texttt{char}

Number of cells born from this cell.

\subsection{Spawned From}

\texttt{char}

Whether this cell spawned from that neighbor.